\label[chap-experiments]
\chap Experiments
This chapter is dedicated to the experimental evaluation of the proposed localization system. Firstly, the description of evaluation approaches is given. As the system is tested in real experiments, the second section is dedicated to implementing the localization system on the robotic platform CART2. This section contains a brief introduction to specific IMU, UWB network and wheel encoders, and the robotic platform itself.

The following two sections are dedicated to experiments done in two different experimental environments. The first set of experiments, described in the third section, were realized in a lab at Datavision s.r.o. company\fnote{To find out more about the company follow link: \url{https://datavision.software/}} with an external localization system based on an AprilTag detection with a camera. This AprilTag localization system was set up specifically for evaluation of localization system proposed in this thesis. Additional experiments took place at the Intelligent and Mobile Robotics lab at the Czech Technical University in Prague - Czech Institute of Informatics, Robotics, and Cybernetics, where the Vicon reference system is used for the evaluation. These experiments are summarized in the fourth section of this chapter.

\sec Description of the evaluation
First, it is necessary to determine which variables exhibit the performance of the system well. Then choose the appropriate test scenarios on which to evaluate the behaviour of the individual variables. It is good to select a suitable reference system and perform the entire evaluation on a set of predefined metrics.

The evaluation is focused on the following states
\begitems
* position x,
* position y
* position in 2D space (xy)
* and angle of rotation (Euler angle yaw).
\enditems

It is worth noting that the output of the proposed system is the position in a 3D space. However, this work aimed to design a localization in a 2D space, and therefore an accent is put on these four states.


For system evaluations, it is advisable to have an external reference system, which will provide us with reference data with which we can compare the results of the evaluated system. The data from the reference and evaluated systems must have unified timestamps to compare their outputs easily.

Experimental scenarios are chosen in the way that their difficulty increases. The first tests are performed in a smaller area. Experiments start with a stationary test and simple movements at shorter distances (moving in one axis, rotating an one spot). Then, more complex trajectories are approached, such as following rectangular and infinity shape trajectories. Further experiments are performed in a larger space where more complex trajectories are tested.

Data from both reference and evaluated systems are then processed as follows. Each reference measurement's timestamp is taken according to which the nearest measurement can be found in the evaluated data. This filtering is vital for reference systems that have a lower frequency than the proposed localization.

The criteria that interest us are
\begitems
* the evolution of individual variables over time,
* the visualization of the trajectories from both systems
* and the errors of variables versus the reference.
\enditems

These metrics are used to analyze whether the proposed system converges or diverges from reality and whether the system is subject to drift. The root mean square (RMSE), average, median, lower and upper quantile, minimum and maximum metrics define the system's resulting accuracy and precision.

Error is usually defined as a difference between reference and reality measurement. It is pretty straightforward for one variable but can be a little bit tricky. The error in position in 2D space $[x, y]$ is defined as the Euclidian distance between reference $[x, y]_{REF}$ and the output of the proposed localization system $[x, y]$ written in the following equation

$$	distance = \sqrt{(x^{REF}-x)^2 + (y^{REF}-y)^2}. \eqmark$$

The root mean square error (RMSE) is defined as

$$	rmse = \sqrt{{\sum_{t=1}^{t=n}(x_t^{REF} - x_t)^2} \over {n}}, \eqmark$$

where
\begitems
* $t$ is the index of the current sample of measurement,
* $n$ is the number of samples in the dataset,
* $x_t^{REF}$ is the current reference measurement
* and $x_t$ is the current measurement.
\enditems

The rmse for euclidian distances then
$$	rmse = \sqrt{{\sum_{t=1}^{t=n}[(x_t^{REF} - x_t)^2 + (y_t^{REF} - y_t)^2]} \over {n}}. \eqmark$$

All these testing scenarios and metrics are further applied in two test environments.

\sec Used hardware description
This section presents an overview of used sensors for the proposed localization and a description of the CART2 robotic platform used during experiments.
\secc Sensors specification
The onboard sensors of interest are the Inertial Measurement Unit, ultra-wideband localization tag, April tag and encoders on motors.

The {\bf inertial measurement unit} used during the experiments is the Epson M-G365PDF1 (loaner sample). The Epson M-G365 is used in various applications ranging from stabilization systems (as a camera gimbal) to navigation systems.

\midinsert
\line{\hsize=.5\hsize \vtop{%
      \clabel[fig-epson-g365]{IMU used in experiments Epson M-G365PDF1}
      \picw=0.5\hsize \cinspic imgs/experiments/g365.jpg
      \caption/f IMU used in experiments Epson M-G365PDF1.
   \par}\vtop{%
      \clabel[fig-epson-avar]{Experimental setup for static data acquisition of Epson M-G365PDF1}
      \picw=\hsize \cinspic imgs/experiments/avar_epson_setup.png
      \caption/f Experimental setup for static data acquisition of Epson M-G365PDF1.
   \par}}
\endinsert

The IMU has six degrees of freedom and measures angular rates and linear accelerations in three axes. It is factory calibrated and the calibration data are stored in the memory of the unit. Technical specifications of the Epson M-G365 can be found at \cite[datasheet_epson], while the summary is included in Table \ref[spec_epson]

\midinsert \clabel[spec_epson]{Technical specifications of Epson M-G365PDF1}
\ctable{lr}{
\hfil
Specification & Value \crl \tskip4pt
Triple gyroscopes & $\pm$ 450 °/sec \cr
Gyroscopes bias instability & 1.2 °/hr \cr
Gyroscopes initial bias error & 0.1 °/s \cr
Angular random walk & 0.08 °/√hr \cr
Tri-axis accelerometers & $\pm$ 10 G \cr
Accelerometers bias instability & 16 $\mu$G \cr
Accelerometers initial bias error & 3 mG \cr
Velocity random walk & 0.033 (m/s)/√hr \cr
}
\caption/t Technical specifications of Epson M-G365PDF1 \cite[epson_g365].
\endinsert

As I already mentioned in Chapter \ref[chap-sec-avar], the AVAR analysis of IMU sensors can give us a brief overview of IMU's specifications. The experimental setup for static data acquisition can be seen in Figure \ref[fig-epson-avar]. The sensor is mounted on two sponges and fixed with cardboard. The static data were recorded for 48 hours at a frequency of 30.0 Hz. The experiment took place in a village without any subways, trams or trains to reduce the external vibrations (to reduce potential outliers) and at standard room temperature (about 23 °C).

For the AVAR computation, I used a python library named AllanTools. For the purposes of this thesis, the overlapping Allan deviation function is used.

\picw=.6\hsize
\centerline {\inspic imgs/experiments/epson_accelerometers.pdf \hfil \inspic imgs/experiments/epson_gyroscopes.pdf }\nobreak
\centerline {a) AVAR for accelerometers\hfil\hfil b)AVAR for gyroscopes}\nobreak\medskip
\clabel[fig-epson-avar2]{Overlapping Allan variance plot for Epson M-G365PDF1.}
\caption/f Overlapping Allan variance plot for Epson M-G365PDF1.

As shown in Figure \ref[fig-epson-avar2], the external reset for the integration of IMU measurement should be performed at least every 100 seconds for the results not to be corrupted by the non-Gaussian noise. The UWB localization system works at 10 Hz, which should be frequent enough.

The IMU is set up for the final experiments to publish the delta angle and the delta velocity at 100 Hz with a moving average filter with tap 64. Because of that, higher frequencies above 10 Hz are filtered, as shown in Figure \ref[fig-epson-filter].

\midinsert
    \picw=0.9\hsize \cinspic imgs/experiments/epson_filter.png
    \clabel[fig-epson-filter]{Moving average filter characteristics for Epson M-G365PDF1.}
    \caption/f Moving average filter characteristics for Epson M-G365PDF1 \cite[epson_g365].
\endinsert

The {\bf UWB localization system} is provided by Qorvo's MDEK1001 ultra-wideband development kit. This kit includes 12 DWM1001-DEV development boards completely enclosed in plastic, see Figure \ref[fig-uwb-dwm].

\midinsert
    \picw=0.4\hsize \cinspic imgs/experiments/mdek1001.jpg
    \clabel[fig-uwb-dwm]{DWM1001-DEV development boards.}
    \caption/f DWM1001-DEV development boards \cite[mdek_1001].
\endinsert

Each board can be configured as an anchor, tag or bridge node. The system is installed with six fixed anchors that are mounted and one tag which can move. The anchors are higher than the tag. In the network, the initiator anchor provides the synchronization of transmitted data. The example setup of the system can be seen in Figure \ref[fig-uwb-anchors_tags].

\midinsert
    \picw=0.8\hsize \cinspic imgs/experiments/uwb_anchors_tags.png
    \clabel[fig-uwb-anchors_tags]{Positioning of anchors and tags.}
    \caption/f Positioning of UWB anchors and tags \cite[mdek_1001].
\endinsert

The anchors should all be mounted at the same height and higher than the operating area of the moving tag. They also should not be mounted close to any metal to get the best accuracy possible. There is a mobile application available to configure the network. The positions of the anchors are estimated manually and set in the network configuration.

Technical specifications of the MDEK1001 and the DWM1001-DEV can be found in documents \cite[mdek_1001], the summary is listed in Table \ref[spec_mdek1001].

\midinsert \clabel[spec_mdek1001]{System performance of MDEK1001}
\ctable{lr}{
\hfil
Specification & Value \crl \tskip4pt
Localization technology & Two-way ranging \cr
Maximum tag location rate & 10 Hz \cr
X-Y location accuracy & < 10 cm \cr
Point to point range & up to 60 m in line of sight conditions \cr
Scheme range & 25 - 30 m between anchors\cr
}
\caption/t System performance of MDEK1001 \cite[mdek_1001].
\endinsert

The {\bf odometry} is computed according to the measurement of encoders in Maxon EPOS4 positioning controllers for Maxon brushless DC motors.

{\bf AprilTag} serves as a global reference for the pose of the CART2 platform. The AprilTag detection software computes the precise 3D position, orientation, and identity of the tags relative to the camera \cite[april_tag]. This tag is similar to QR codes (a type of two-dimensional bar code), but it encodes smaller data payloads (between 4-12 bits), and it can be detected more robustly.

The {\bf camera} used to detect the AprilTag was the Niceboy Stream Pro with Full HD (1920 x 1080) resolution, 30 FPS, 90 ° field of vision and a f/1.8 lens aperture \cite[camera_niceboy].

\midinsert
\line{\hsize=.5\hsize \vtop{%
      \clabel[fig-apriltag]{AprilTag used for detection of CART2 position}
      \picw=4cm \cinspic imgs/experiments/apriltag.png
      \caption/f AprilTag used for detection of CART2 position at Datavision s.r.o.
   \par}\vtop{%
      \clabel[fig-camera]{Camera used for AprilTag detection Niceboy Stream Pro}
      \picw=4cm \cinspic imgs/experiments/camera.jpg
      \caption/f Camera used for AprilTag detection Niceboy Stream Pro \cite[camera_niceboy].
   \par}}
\endinsert

\secc CART2 platform description and sensors placements
An image of the utilized CART2 platform can be seen in Figure \ref[fig-cart2_close]. The coordinate frame of CART2, called baselink,  is illustrated in all Figures \ref[fig-cart2_close] and \ref[fig-cart2_sensors]. The CART2 used for various robotic competitions is a differential drive equipped with the ADlink MXE-210 computer.

\midinsert
\centerline {\picw=4cm \inspic imgs/experiments/cart2_up.pdf \hfil\hfil \picw=7cm \inspic imgs/experiments/cart2_side.pdf }\nobreak
\centerline {a) Top view \hfil\hfil b) Side view}\nobreak\medskip
\clabel[fig-cart2_close]{A photo of the utilized CART2 platform}
\caption/f A photo of the utilized CART2 platform.
\endinsert

The onboard sensor placements are illustrated in Figure \ref[fig-cart2_sensors]. The vehicle uses a Maxon brushless DC motor (Maxon EC-imotor) controlled by Maxon EPOS4 control units.

\midinsert
\line{\hsize=.5\hsize \vtop{%
      \clabel[fig-cart2_sensors]{A photo of sensors placements on CART2}
      \picw=6cm  \cinspic imgs/experiments/cart2_whole.pdf
      \caption/f A photo of sensors placements on CART2.
   \par}\vtop{%
      \clabel[fig-cart2_coordinates]{Illustration of coordinate systems at CART2}
      \picw=\hsize \cinspic imgs/experiments/cart2_coordinates.pdf
      \caption/f Illustration of coordinate systems at CART2.
   \par}}
\endinsert

The UWB tag is mounted on a wooden stick approximately one meter above the CART2 platform to reduce any reflections of the UWB waves from surfaces and the negative influence of any metal parts. The AprilTag is mounted below the UWB tag not to become a barrier in the UWB wave. When the AprilTag is mounted on top of the CART2 platform, the wooden stick obstructs the camera from detecting the AprilTag. Thus, I decided to mount it as high as possible to reduce any of these situations from happening.

{\bf Coordinate systems of CART2 and the placement of the sensors} are illustrated in Figure \ref[fig-cart2_coordinates] and are described in Table \ref[cart2_frames]. Note that the world, the camera and the uwb_base frames are fixed within the environment and are explained in Section \ref[sec-experiments_bu04]. This work aims to localize the position and orientation of the baselink, i.e., to estimate the position of the baselink expressed in the world frame and the orientation of the baselink with respect to the world frame.

\midinsert \clabel[cart2_frames]{Summary of transforms for experiments with CART2}
\ctable{lrr}{
\hfil Transform & Translation [x, y, z] in [m] & Rotation in quaternion [x, y, z, w] \crl \tskip4pt
uwb_tag to baselink & [0.0, 0.0, 1.192] & [0.0, 0.0, 0.0, 1.0] \cr
april_tag to baselink & [0.0, -0.171, 1.071] & [0.707107, -0.707107, 0.0, 0.0] \cr
imu to baselink & [0.0, 0.0, 0.05] & [0.0, 0.0, -0.7071068, 0.7071068] \cr
}
\caption/t Summary of transforms for experiments with CART2.
\endinsert

\label[sec-experiments_bu04]
\sec Experiments in the lab at Datavision s.r.o.
\secc Experimental lab description
	For the first experiments, I created an experimental setup at Datavision s.r.o. with the global reference given by the camera detection from the AprilTag mounted on the CART2 platform. The address of the building is Ukrajinská 1487/2a, 101 00 Prague 10 - Vršovice. The dimensions of the room for the experiments are approximately 4 x 6 [m], and the camera view area is approximately 2.5 x 4.5 [m]. The setup can be seen in Figure \ref[fig-bu04].
	
\midinsert
    \picw=\hsize \cinspic imgs/experiments/bu04_coordinates.pdf
    \clabel[fig-bu04]{Experimental setup at Datavision s.r.o.}
    \caption/f Experimental setup at Datavision s.r.o..
\endinsert

Firstly, UWB anchors need to be mounted and measure their poses according to the world coordinate system. These poses need to be set in the mobile application for the configuration of the UWB network. The world frame coincides with the uwb_base frame in rotation and only differs in z coordinate in translation.

\midinsert \clabel[bu04_frames]{Summary of transforms for experiments setup at Datavision s.r.o.}
\ctable{lrr}{
\hfil Transform & Translation [x, y, z] in [m] & Rotation in quaternion [x, y, z, w] \crl \tskip4pt
world to uwb_base & [0.0, 0.0, 2.58] & [0.0, 0.0, 0.0, 1.0] \cr
world to camera & [1.089, 2.024, 2.625] & [0.707107, -0.707107, 0.0, 0.0] \cr
}
\caption/t Summary of transforms for experiments setup at Datavision s.r.o.
\endinsert

Secondly, the camera needs to be mounted and measure its position according to the world coordinate system. The transformation between the world and the camera is not trivial. Because of that, I decided to use the transformation mentioned in Table \ref[bu04_frames] and then computed a camera homography according to a few positions measured by camera detection and also by hand. The homography defines the transformation between a planar surface (ground) and a camera image plane. The camera homography is then applied to the AprilTag detection pose, which is considered a global reference. These frames are illustrated in \ref[fig-cart2_coordinates].
	
\secc Description of experiments
	The area for experiments is not big. The distance between anchors is 2.5 or 5.0 m, and such setup with nearby anchors significantly downgrades the performance of the UWB localization system. However, I decided that these experiments serve as a proof of concept of the localization idea.
	
\midinsert
\centerline {\picw=4.75cm \inspic imgs/experiments/trajectories/stat_and_rot_move.pdf \hfil\hfil \picw=5cm \inspic imgs/experiments/trajectories/xy_move.pdf }\nobreak
\centerline {a) Stationary and rotation tests \hfil\hfil b) X and Y tests}\nobreak\medskip
\centerline {\picw=5cm \inspic imgs/experiments/trajectories/rectangle.pdf \hfil\hfil \picw=4.7cm \inspic imgs/experiments/trajectories/infinity.pdf }\nobreak
\centerline {c) Rectangle test \hfil\hfil d) Infinity test}\nobreak\medskip
\clabel[fig-bu04-traj]{Trajectories for experiments at Datavision s.r.o.}
\caption/f Trajectories for experiments at Datavision s.r.o..
\endinsert
	
Experiments can be divided based on six simple trajectories of CART2 into
 \begitems
 	* stationary test (stationary test),
 	* rotation above 360 degrees in one direction at one place test (rotation test),
 	* moving in x direction (x test),
 	* moving in y direction (y test),
 	* moving in rectangle shape (rectangle test)
 	* and moving in infinity shape (infinity test).
 \enditems
 I picked a specific initial pose for each testing trajectory to repeatedly make these tests in very similar conditions. Also, rotation, rectangle, and infinity tests ran several times in a row without start/stop of the system to see how it behaves in a long term.
 
 Trajectories of these tests and the starting and ending positions of CART2 performing the movement are illustrated in Figure \ref[fig-bu04-traj]. CART2 platform was controlled via keyboard and joystick during experiments. The controlling via keyboard has the benefit of control velocity in each direction easily. Thus I used it for constant speed during simple moving in single-axis and simple rotation around a single axis. With that, I controlled all tests, except the infinity test, where the movement is complex. During the rectangle test, CART2 drove forward (moving in single-axis) then stopped and turned at one place above 90 degrees (rotating above single axis). I controlled the CART2 platform with the joystick in infinity shape movement and tried to move similar velocities as in constant movements. The summary of velocities during experiments is given in Table \ref[bu04_traj_vel].
 
\midinsert \clabel[bu04_traj_vel]{Velocities during experiments in Datavision s.r.o.}
\ctable{lrrr}{
\hfil Test & Description & Speed [m/s] & Turn [rad/s] \crl \tskip4pt
stationary & Constant velocities & 0.0 & 0.0 \cr
rotation & Constant velocities & 0.0 & 0.1094 \cr
x & Constant velocities & 0.0750 & 0.0 \cr
y & Constant velocities & 0.0750 & 0.0 \cr
rectangle & Constant velocities & 0.0750 or 0.0 & 0.0 or 0.1094 \cr
infinity & non constant (control via joystick) & approx. 0.0750 & approx. 0.1094 \cr
}
\caption/t Velocities during experiments in Datavision s.r.o..
\endinsert
  
\secc Evaluation of experiments
Most of the experiments were performed in this environment. However, I select only the most interesting ones for the evaluation in the thesis. The most attractive experiments are stationary, rectangle and infinity shape tests, which are presented bellow.


{\bf Stationary test}


During the stationary test, the CART2 is not moving, and the experiment lasts about 16 minutes. As shown in Figure \ref[fig-bu04-static], the precision of the position does not change significantly over time. The position in the proposed system is determined mainly by the absolute localization of the UWB, but it can be seen that our system can further refine the UWB localization. There is a significant drift in the orientation angle yaw caused by the INS.  The orientation error is corrected by the linear velocities given by odometry measurement, but because the CART2 is not moving, the error is not corrected, and the drift slowly grows.
\midinsert
\centerline {\picw=8cm \inspic imgs/experiments/bu04/static/position_xy.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/bu04/static/orientation_yaw.pdf }\nobreak
\centerline {a) Position in 2D space \hfil\hfil b) Yaw angle in time}\nobreak\medskip
\centerline {\picw=8cm \inspic imgs/experiments/bu04/static/position_x.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/bu04/static/position_y.pdf }\nobreak
\centerline {c) Position x in time \hfil\hfil d) Position y in time}\nobreak\medskip
\clabel[fig-bu04-static]{Stationary test, evaluation of trajectory at Datavision s.r.o.}
\caption/f Stationary test, evaluation of trajectory at Datavision s.r.o.
\endinsert

In general, box plots show the five-number summary of a set of data: including the minimum, first (lower) quartile, median (interquantile), third (upper) quartile, and maximum. I added a visualisation of the RMSE, to keep these information together.

In Figures \ref[fig-bu04-static-box_pos] and \ref[fig-bu04-static-box_ori] are all these metrics for the position in 2D error and the orientation error during the stationary test shown. Sufficient information from Figure \ref[fig-bu04-static-box_ori] is that maximum orientation error during 16 minutes long stationary test is not bigger than ten degrees. The position error is small and does not increase in Figure \ref[fig-bu04-static-box_pos]. It is worth noticing that the external reference frequency is not high (1-2 Hz). When filtering the corresponding measurement samples for external reference timestamp, closer values may be favoured if the localization fluctuates. Thus, there may be a slight deformity here.

\midinsert
\line{\hsize=.5\hsize \vtop{%
      \clabel[fig-bu04-static-box_pos]{Stationary test: boxplot, rmse, min and max analysis of the position 2D error, at Datavision s.r.o.}
      \picw=\hsize \cinspic imgs/experiments/bu04/static/odometry_position_box_plot.pdf
      \caption/f Stationary test: boxplot, rmse, min and max analysis of the position 2D error, at Datavision s.r.o.
   \par}\vtop{%
      \clabel[fig-bu04-static-box_ori]{Stationary test: boxplot, rmse, min and max analysis of the orientation error, at Datavision s.r.o.}
      \picw=\hsize \cinspic imgs/experiments/bu04/static/orientation_box_plot.pdf
      \caption/f Stationary test: boxplot, rmse, min and max analysis of the orientation error, at Datavision s.r.o.
   \par}}
\endinsert


{\bf Rectangle-shaped test}

\midinsert
\centerline {\picw=8cm \inspic imgs/experiments/bu04/rectangle/position_xy.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/bu04/rectangle/orientation_yaw.pdf }\nobreak
\centerline {a) Position in 2D space \hfil\hfil b) Yaw angle in time}\nobreak\medskip
\centerline {\picw=8cm \inspic imgs/experiments/bu04/rectangle/position_x.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/bu04/rectangle/position_y.pdf }\nobreak
\centerline {c) Position x in time \hfil\hfil d) Position y in time}\nobreak\medskip
\clabel[fig-bu04-rectangle]{Rectangle test, analysis of trajectory, Datavision s.r.o.}
\caption/f Rectangle-shaped test, analysis of trajectory, Datavision s.r.o.
\endinsert

During the experiment, nine repetitions of the rectangle-shaped trajectory were performed to define the localization's accuracy, precision, and repeatability. The course of one such passage is shown in Figure \ref[fig-bu04-rectangle]. The waves in the left part of Figure \ref[fig-bu04-rectangle]a can be caused by reflections of the UWB signal as CART2 moves close to different furniture and boxes. This phenomenon can be seen in all reproductions of testing scenarios. However, despite this fact, the system manages to locate CART2 with sufficient precision. The static offset at the bottom of Graph \ref[fig-bu04-rectangle]a is possibly given by a short line of sight between the localization anchors or the transformation inaccuracies in the external reference.

\midinsert
\line{\hsize=.5\hsize \vtop{%
      \clabel[fig-bu04-rectangle-box_pos]{Rectangle-shaped tests: boxplot, rmse, min and max analysis of the position 2D error, at Datavision s.r.o.}
      \picw=\hsize \cinspic imgs/experiments/bu04/rectangle/odometry_position_box_plot.pdf
      \caption/f Rectangle-shaped tests: boxplot, rmse, min and max analysis of the position 2D error, at Datavision s.r.o.
   \par}\vtop{%
      \clabel[fig-bu04-rectangle-box_ori]{Rectangle-shaped test: boxplot, rmse, min and max analysis of the orientation error, at Datavision s.r.o.}
      \picw=\hsize \cinspic imgs/experiments/bu04/rectangle/orientation_box_plot.pdf
      \caption/f Rectangle-shaped test: boxplot, rmse, min and max analysis of the orientation error, at Datavision s.r.o.
   \par}}
\endinsert

Figures \ref[fig-bu04-rectangle-box_pos] and \ref[fig-bu04-rectangle-box_ori] show that the localization results are well repeatable. RMSE in position is held at 7 cm and in the orientation is not higher than 3 degrees. The maximum errors in the position are about 15 cm and in the orientation 8 degrees. More than 50 \% of the errors are kept below 10 cm in the position and 3 degrees in the orientation.

{\bf Infinity-shape test}
\midinsert
\centerline {\picw=8cm \inspic imgs/experiments/bu04/infinity/position_xy.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/bu04/infinity/orientation_yaw.pdf }\nobreak
\centerline {a) Position in 2D space \hfil\hfil b) Yaw angle in time}\nobreak\medskip
\centerline {\picw=8cm \inspic imgs/experiments/bu04/infinity/position_x.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/bu04/infinity/position_y.pdf }\nobreak
\centerline {c) Position x in time \hfil\hfil d) Position y in time}\nobreak\medskip
\clabel[fig-bu04-infinity]{Infinity test, analysis of trajectory, Datavision s.r.o.}
\caption/f Infinity-shaped test, analysis of trajectory, Datavision s.r.o.
\endinsert

Evaluating the infinity-shaped trajectory brings the most complex view because the CART2 here moves simultaneously in linear and angular velocity (Figure \ref[fig-bu04-infinity]). Ten repetitions of such a trajectory demonstrate how the system behaves in the most realistic scenario.

\midinsert
\line{\hsize=.5\hsize \vtop{%
      \clabel[fig-bu04-infinity-box_pos]{Infinity-shaped: boxplot, rmse, min and max analysis of the position 2D error, at Datavision s.r.o.}
      \picw=\hsize \cinspic imgs/experiments/bu04/infinity/odometry_position_box_plot.pdf
      \caption/f Infinity-shaped: boxplot, rmse, min and max analysis of the position 2D error, at Datavision s.r.o.
   \par}\vtop{%
      \clabel[fig-bu04-infinity-box_ori]{Infinity-shaped: boxplot, rmse, min and max analysis of the orientation error, at Datavision s.r.o.}
      \picw=\hsize \cinspic imgs/experiments/bu04/infinity/orientation_box_plot.pdf
      \caption/f Infinity-shaped: boxplot, rmse, min and max analysis of the orientation error, at Datavision s.r.o.
   \par}}
\endinsert
As can be seen in Figures \ref[fig-bu04-infinity-box_pos] and \ref[fig-bu04-infinity-box_ori], the errors are small. RMSE of position is lower than 10 cm, and RMSE of orientation is below 2 degrees. There are maximum errors of 16 cm in position and 7 degrees in orientation, but most do not exceed 10 cm and 1 degree, respectively.


Notice that the precision of the proposed localization system meets the limits of the precision of the external reference system.

\sec Experiments at CIIRC
\secc Experiments at CIIRC description
\midinsert
\centerline {\picw=0.85\hsize \inspic imgs/experiments/ciirc1.pdf }\nobreak
\centerline {a) View from camera 1}\nobreak\medskip
\centerline {\picw=0.85\hsize \inspic imgs/experiments/ciirc2.pdf }\nobreak
\centerline {b) View from camera 2}\nobreak\medskip
\clabel[fig-ciirc]{Experimental setup at the Intelligent and Mobile Robotics lab (CIIRC)}
\caption/f Experimental setup at the Intelligent and Mobile Robotics lab (CIIRC).
\endinsert
The next experimental environment was placed at the Czech Technical University in Prague – Czech Institute of Informatics, Robotics, and Cybernetics (CIIRC) in the Intelligent and Mobile Robotics lab. The localization system was evaluated using the VICON external camera localization system \cite[vicon]. The VICON defines the world frame. UWB localization is installed with six anchors and their positions are defined in a world frame. The setup is illustrated in Figure \ref[fig-ciirc]. The benefit of this environment is the size of the laboratory. The minimal distance between two anchors is larger than 3,5 meters; thus, the accuracy of the UWB localization should be increased.

Experiments here are more complex because they are longer both in terms of duration or distance. These experiments promised to evaluate possible drift caused by using two relative localizations based on IMU and odometry. As I already mentioned in previous chapters, IMU’s drift increases over time and odometry’s drift grow larger over distance travelled. Performed experiments can be divided based on three simple trajectories of CART2 into
 \begitems
 	* moving in the shape of a rectangle (rectangle test),
 	* moving in the shape of an infinity symbol (infinity test)
 	* and so-called joy ride, where CART2 tries to move as long-distance as possible without crossing its trajectory (joy ride test).
 \enditems
 I picked a specific initial pose for each testing trajectory to repeatedly make these tests in very similar conditions. Also, the rotation, the rectangle and the infinity tests ran several times in a row without starting or stopping the system to see how it behaves under longer time durations.

\secc Evaluation of experiments at CIIRC

{\bf Rectangle-shaped test}


In Figure \ref[fig-ciirc-rectangle], there is no major problem with estimating the position, which is close to the external reference even at greater distances travelled. There is no drift in the position on a longer trajectory, and the filter continuously converges to the correct solution. It can be seen that the position estimation accuracy is closely related to the positioning accuracy from the UWB localization, and thus when CART2 passes very close to the initiator anchor, the position estimation get worse. Unfortunately, in all experiments, there was a problem with determining the orientation. The results do not very much correspond to the results of the experiments in the Datavision s.r.o. Laboratory, I propose that the problem may have occurred during the actual measurement in the IMR CIIRC during the setup of the CART2 platform. But unfortunately, I was not able to analyze it properly by the deadline for submitting this thesis.

\midinsert
\centerline {\picw=8cm \inspic imgs/experiments/ciirc/rectangle/position_xy.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/ciirc/rectangle/orientation_yaw.pdf }\nobreak
\centerline {a) Position in 2D space \hfil\hfil b) Yaw angle in time}\nobreak\medskip
\centerline {\picw=8cm \inspic imgs/experiments/ciirc/rectangle/position_x.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/ciirc/rectangle/position_y.pdf }\nobreak
\centerline {c) Position x in time \hfil\hfil d) Position y in time}\nobreak\medskip
\clabel[fig-ciirc-rectangle]{Rectangle-shaped test, analysis of trajectory, at IMR CIIRC}
\caption/f Rectangle-shaped test, analysis of trajectory, at IMR CIIRC
\endinsert

The experiment was repeated a total of three times. Analysis of error is given in Figure \ref[fig-ciirc-rec-box]. At the beginning of each iteration, the system rebooted. The system's errors turn out to be
\begitems
*repeatable, as the error values do not differ significantly from each other,
*and accurate, as the RMSE was less than 10 cm for all three passes.
\enditems
The maximum values here could exceed 40 cm. However, these extremes are caused by outliers in the external reference.

\midinsert
    \picw=0.6\hsize \cinspic imgs/experiments/ciirc/rectangle/multi_odometry_position_box_plot.pdf
    \clabel[fig-ciirc-rec-box]{Rectangle-shaped tests: boxplot, rmse, min and max analysis of the position error, at IMR CIIRC}
    \caption/f Rectangle-shaped tests: boxplot, rmse, min and max analysis of the position error, at IMR CIIRC
\endinsert

Another interesting experiment was to go through several iterations of the same rectangle-shaped trajectory without shutting down the system. As can be seen in the Figure \ref[fig-ciirc-rep_rec], the individual iteration does not differ from each other. It can therefore be seen that there is no significant drift in the position estimation.

\midinsert
\centerline {\picw=8cm \inspic imgs/experiments/ciirc/rep_rect/position_xy.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/ciirc/rep_rect/orientation_yaw.pdf }\nobreak
\centerline {a) Position in 2D space \hfil\hfil b) Yaw angle in time}\nobreak\medskip
\centerline {\picw=8cm \inspic imgs/experiments/ciirc/rep_rect/position_x.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/ciirc/rep_rect/position_y.pdf }\nobreak
\centerline {c) Position x in time \hfil\hfil d) Position y in time}\nobreak\medskip
\clabel[fig-ciirc-rep_rec]{Multiple rectangle-shaped test without shut down, analysis of trajectory, at IMR CIIRC}
\caption/f Multiple rectangle-shaped test without shut down, analysis of trajectory, at IMR CIIRC
\endinsert


{\bf Infinity-shaped test}


During the infinity-shaped test in Figure \ref[fig-ciirc-infinity], the position estimation is also sufficient. However, the detection from the global reference sometimes failed and created a few outliers. Thus, the results are a bit skewed.

\midinsert
\centerline {\picw=8cm \inspic imgs/experiments/ciirc/infinity/position_xy.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/ciirc/infinity/orientation_yaw.pdf }\nobreak
\centerline {a) Position in 2D space \hfil\hfil b) Yaw angle in time}\nobreak\medskip
\centerline {\picw=8cm \inspic imgs/experiments/ciirc/infinity/position_x.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/ciirc/infinity/position_y.pdf }\nobreak
\centerline {c) Position x in time \hfil\hfil d) Position y in time}\nobreak\medskip
\clabel[fig-ciirc-infinity]{Infinity-shaped test, analysis of trajectory, at IMR CIIRC}
\caption/f Infinity-shaped test, analysis of trajectory, at IMR CIIRC
\endinsert

{\bf Joy ride test}


As shown in Figure \ref[fig-ciirc-joy], the position estimation worsens if the CART2 manipulates close to the initiator anchor. Also, it is worth to mention, that there is a wide column at [1.8, 1.0] m, which can cause signal reflections.

\midinsert
\centerline {\picw=8cm \inspic imgs/experiments/ciirc/joy/position_xy.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/ciirc/joy/orientation_yaw.pdf }\nobreak
\centerline {a) Position in 2D space \hfil\hfil b) Yaw angle in time}\nobreak\medskip
\centerline {\picw=8cm \inspic imgs/experiments/ciirc/joy/position_x.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/ciirc/joy/position_y.pdf }\nobreak
\centerline {c) Position x in time \hfil\hfil d) Position y in time}\nobreak\medskip
\clabel[fig-ciirc-joy]{Joy ride test, analysis of trajectory, at IMR CIIRC}
\caption/f Joy ride test, analysis of trajectory, at IMR CIIRC
\endinsert

{\bf Sumarize of experiments errors}

\midinsert
\centerline {\picw=8cm \inspic imgs/experiments/ciirc/rectangle/odometry_position_box_plot.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/ciirc/rep_rect/odometry_position_box_plot.pdf }\nobreak
\centerline {a) Single rectangle-shape test \hfil\hfil b) Multiple rectangle-shape test}\nobreak\medskip
\centerline {\picw=8cm \inspic imgs/experiments/ciirc/infinity/odometry_position_box_plot.pdf \hfil\hfil \picw=8cm \inspic imgs/experiments/ciirc/joy/odometry_position_box_plot.pdf }\nobreak
\centerline {c) Infinity-shape \hfil\hfil d) Joy ride test}\nobreak\medskip
\clabel[fig-ciirc-sum]{Summary of errors in the position during experiments at IMR CIIRC}
\caption/f Summary of errors in the position during experiments at IMR CIIRC
\endinsert

Comparison of position errors in individual experiments in Figure \ref[fig-ciirc-sum] shows interesting conclusions. Errors are similar for different test scenarios, not bigger than 30 cm, but usually, RMSE is close to 10 cm. Based on the result of the experiments, the proposed localization system seems viable for indoor localization with high accuracy and precision demands.


